Kindly Bent to Free Us

### 2020-05-30 kindly/extensions.tex
NLL Extension: fixed typo.
Start presenting limitations (and potential extensions)

\lstMakeShortInline[keepspaces,basicstyle=\small\ttfamily]@
\section{Extensions}

\subsection{Algebraic Datatypes}

\subsection{Branching constructs}

\subsection{Non-Lexical Regions}

@@

>>>>

\lstMakeShortInline[keepspaces,basicstyle=\small\ttfamily]@
\section{Limitations and Extensions}

\subsection{Flow sensitivity}

The type system defined so far does not support any
form of flow sensitivity. This means that code patterns that relies on
subtle flow-sensitive usage of permissions and linearity likely will not
typecheck in \lang. As an example the following piece of code,
which defines the merge function on linear lists,
cannot be expressed directly.

\begin{lstlisting}
let rec merge l1 l2 = match l1, l2 with
  | h1::t1, h2::t2 ->
    if &h1 < &h2 
    then h1::(merge t1 l2) (* Must expand l2 to h2::t2 here *)
    else h2::(merge l1 t2)
  | ....
\end{lstlisting}

Such pattern usually requires richer logic, such as
Mezzo~\citep{DBLP:phd/hal/Protzenko14}.
However
\citet{DBLP:journals/corr/abs-1903-00982} formalize Rust's
notion of non-lexical lifetimes which allow to partially support such
code patterns. We believe this notion can be adapted to \lang's notion of
regions

\subsubsection{Non-Lexical Regions}

@@

\subsection{Capabilities and Identity}

One primary feature of \lang is that its tracking of linearity does
not use any notion of ``identity'': the type system can't specify that two
objects are the same, simply that they share the same usage pattern with
regards to linearity.
Languages such as Alms~\citep{DBLP:conf/popl/TovP11}, on the other hand,
often relies on a notion of identity to express capabilities.
For instance, the Alms function
@Array.create@ of type @int -> 'a -> \E 'b. ('a, 'b) array@ uses
@'b@ to uniquely identifies the array.
Functions such as @Array.acquire : ('a, 'b) array -> 'b cap@
are then used to manipulate capabilities pertaining to the array.

While such uses are partially covered by borrows and regions,
a notion of identity associated to regions
would allow expressing regions directly in type signatures.
For instance, the @get_eb@ function shown in
\cref{tuto:vector} could be made safe
by creating a restricted inner region on function application,
with the signature:
@&!('k, 'a Array.t) -> int -> \E ('k' < 'k) &!('k', 'a)@

This notion of identity often relies on existential types.
\lang does not support existentials so far as it would
forgo principal type inference.
However existentials
are compatible with the HM(X) framework~\citep{DBLP:conf/icfp/Simonet03}
and would make a very desirable addition to \lang. OCaml also show
that existential types can be made usable without compromising
inference in the rest of the language, by integrating unpacking
and pattern matching.

\subsection{Traditional functional features}

\subsubsection{Algebraic Datatypes}

\subsubsection{Branching constructs}
